### 2020-02-21 sections/srfm.tex
added subsections

\section{Super-Resolved Fluorescence Microscopy}
\label{sec:srfm}
\subsection{General approach}

\subsection{Camera-based techniques}

\subsection{Scanning techniques}
%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../manuscript"
%%% End: