### 2020-10-09 sections/introduction.tex
added caption for resolution figure

\section{Introduction}
\label{sec:introduction}
Many cellular processes like exo- and endocytosis, cell growth, division and
migration, to name only a few, require the rearrangement, growth or shrinkage
of the cellular membrane. Scanning Ion Conductance Microscopy (SICM)
\cite{Hansma1989} is a contact-free scanning probe technique which, if applied
to cellular specimen, provides information about the dynamics of the
topography of the cell membrane. However, the underlying dynamics of the
molecular machinery which orchestrates the cell membrane dynamics remain
unresolved by SICM investigations. The latter can be resolved by fluorescence
microscopy (FM), which allows tagging a protein of interest with a fluorescent
marker and observing its dynamics. Therefore, correlating fluorescence and
scanning ion conductance microscopy data might help to unravel the mechanisms
that underlie the cellular processes listed above that occur at the cellular
membrane.

\begin{figure}
  \sidecaption[t]
  \includegraphics{gr/intro/Fig_SICM-resolution}
  \caption{\textbf{Resolution of SICM.} \textbf{a:} Height profiles (plotted
    vertically) of modelled SICM recordings of two cilindrical particles
    (height and diameter: $r_\text{i}$) at varying distances. \textbf{b:} Two
    consecutive SICM recordings of the equatorial segment of a living boar
    spermatozoon recorded 10 minutes apart with a scanning pipette with
    $r_\mathrm{i} \approx 6.75\,\text{nm}$. \textbf{a:} reprinted with
    permission from ref. \cite{rheinlaender:094905}, \textbf{b:} reprinted
    with permission from ref. \cite{Shevchuk2006}. 
  }
  \label{fig:sicm-resolution}
\end{figure}

The resolution of SICM is determined by the inner opening radius
$r_\mathrm{i}$ of the scanning pipette and can be approximated as
$3r_\mathrm{i}$ (Fig.~\ref{fig:sicm-resolution}a)
\cite{rheinlaender:094905,Rheinlaender2015}.  Thus, the resolution of SICM is
only limited by the technical limits to produce pipettes with very small
openings. Pipettes with radii of approximately 6.75\,nm have been used in for
SICM imaging (Fig.~\ref{fig:sicm-resolution}b) \cite{Shevchuk2006}. In
contrast, the resolution in light microscopy is fundamentally limited by the
diffraction of light. As Ernst Abbe found out in 1873, a beam of light cannot
be focused to a spot smaller than the resolution limit $d_\text{min}$
\cite{Abbe1873}:
\begin{equation}
  d_\text{min} = \frac\lambda{2n\sin\alpha} = \frac\lambda{2\mathrm{NA}}\text{.}
  \label{eq:diffraction-limit}
\end{equation}
Here, $\lambda$ denotes the wavelength of the light used to generate the
image, $n$ the diffractive index of the medium between sample and objective
lense and $\alpha$ the half-cone opening angle of the objective lense. The
product $n\sin\alpha$ is called the numerical aperture $\mathrm{NA}$ of the
objective. The exact limit, of course, depends on $\lambda$ and the NA of the
objective, but as a rule of thumb, it is often stated that the resolution
limit is in the range of 200\,nm. This is approximately 14 times the size of
the smallest structures observed in the SICM recordings shown in
Fig.~\ref{fig:sicm-resolution}B, which were reported to be 14\,nm
\cite{Shevchuk2006}.

\begin{figure}
  \sidecaption
  \includegraphics{gr/intro/Fig_SICM-and-Light-resolution}%
  \caption{\textbf{Hypothetical fluorescence analysis of proteins in the boar
      spermatozoon.} \textbf{a:} Proteins A and B in the stable structures
    observed by SICM. \textbf{b:} Hypothetical positions of the fluorescent
    labels of proteins A and B. Scale bar: 200\,nm \textbf{c:} Computed
    fluorescence image, resolutions were 250\,nm for the magenta and 300\,nm
    for the green marker.}
  \label{fig:sicm-and-light-resolution}
\end{figure}
In all stable structures in the recordings (highlighted by the dashed lines)
at least one group of three protrusions arranged in a line are observable. To
exemplify the impact of the limited resolution of light microscopy, assume,
without any biological support, that such a complex is comprised of two
different proteins A and B, which float in the plasma membrane until they form
the complex. In the complex, protein A is located at the first and last
position (magenta arrows in Fig.~\ref{fig:sicm-and-light-resolution}A), and
protein B on the central spot (green arrows in
Fig.~\ref{fig:sicm-and-light-resolution}a). Further, assume that the two
different proteins have been labelled by two different fluorescent markers
(the positions of the markers are indicated by the magenta and green dots in
Fig.~\ref{fig:sicm-and-light-resolution}b) and a correlating light microscopic
image has been recorded (Fig.~\ref{fig:sicm-and-light-resolution}c) to analyze
the molecular composition of the stable structures.

Due to the limited resolution, the fluorescence image consists of overlapping
spots of the two different markers and looks blurred, compared to the
SICM topography image. Although co-localisation of the two proteins can be
observed at the positions of the stable structures, it is hard if not
impossible to allocate the fluorescent spots to protrusions in the topography
image.

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../manuscript"
%%% End: