### 2020-10-13 sections/correlated-sicm-and-fm.tex
commented the nanoinjection section. Is it really necessary?

\section{Correlating SICM and Fluorescence Microscopy}
\label{sec:correlating-sicm-and-fm}

% Assume a figure here which shows the different degrees of correlation
\begin{figure}
  \includegraphics[width=\textwidth]{gr/correlating/sicm+fm}
  \caption{%
    \textbf{Combining SICM and wide-field fluorescence microscopy.}
    \textbf{a:}
    \textbf{b:} FM image of presynaptic nerve terminals stained with
    FM1-43. \textbf{c:} Overlay of the FM-image and two SICM recordings of the
    same region. \textbf{d:} 3D-representation of the lower right SICM
    recording from c.
    \textbf{b--d:} Reprinted with permission from ref \cite{Scheenen2015}.
  }
  \label{fig:sicm+fm}
\end{figure}

Correlative SICM and FM, despite the mismatch in resolution, allows to gather
information which cannot be obtained by using a single technique
alone. However, the degree of correlation as well as the technical
implementation differs, depending on the FM technique combined with SICM, as
we will detail in the following sections.

Although recording SICM and FM data independently on different samples under
the same conditions can support conclusions that could not be drawn from data
from a single technique alone \cite{Gesper2017,Lee2013,Lyon2009}, we will
focus on applications were SICM and FM data has been recorded from the same
sample. Furthermore, we focus on far-field FM techniques and do not review the
combination of SICM with nearfield optical microscopy  
\cite{Korchev2000,Shevchuk2001,Rothery2003,Bruckbauer2002}.

\subsection{Combining SICM and wide-field FM}
\label{sec:SICM+widefield}
Combining SICM and wide-field FM allows to image the same region of the sample
with both techniques (Fig.~\ref{fig:sicm+fm}a). Due to the different fields of
view, the different pixel sizes and the different times required to record one
image, the two data sets are not recorded synchronously and thus cannot by
correlated pixel-by-pixel. Technically, the combination is simple, since
the SICM only has to be mounted onto an inverted optical FM and both
instruments can be controlled independently by their own software.

So far, the combination of SICM and wide-field FM has been used to find a
region of interest such as an area containing active synapses in a culture of
a neuronal network \cite{Scheenen2015} (Fig.~\ref{fig:sicm+fm}b--d) or to
identify the neuron which had been investigated by SICM-guided patch-clamp of
synaptic boutons \cite{Novak2013}.

Furthermore, it has been used to first identify primary cilia in a cell
culture and subsequently use SICM to reveal whether the selected cilium was
extending into the extracellular space or was embedded in the cell, an
information which could not be retrieved by using FM alone \cite{Zhou2018}.

\begin{figure}
  \includegraphics[width=\textwidth]{gr/correlating/Nikolaev2010-F2}
  \caption{
    \textbf{a:}\textbf{b:}\textbf{c:}\textbf{d:}\textbf{e:}\textbf{f:}
  }
  \label{fig:nikolaev2010}
\end{figure}

Cardio-myocytes exhibit a unique topography with transverse (T-)tubules and
crests (Fig.~\ref{fig:nikolaev2010}a, b). By first recording the topography
and subsequently moving the SICM pipette to either a T-tubule or a crest,
followed by the local application of agonists of either the $\beta_1$- or
$\beta_2$-adrenergic receptor (AR) and recording the intensity ratio of cyan
and yellow fluorescent protein, coupled to a Förster resonance energy transfer
(FRET) based reporter of cyclic AMP (Fig.~\ref{fig:nikolaev2010}c--f), it was
shown that the distribution of the $\beta_2$-AR was altered in failing
cardio-myocytes \cite{Nikolaev2010}. Furthermore, this approach was used to
investigate the effect of partial mechanical unloading, showing that it
reduces the cyclic AMP response of $\beta_2$-ARs at the T-tubules
\cite{Wright2018}. To investigate the distribution of the $\beta_3$-AR, the
approach was extended to use a FRET-based reporter of cyclic GMP, showing that
a redistribution of the $\beta_3$-AR occurs in failing cardio-myocytes, too
\cite{Schobesberger2020}.

\subsection{Combining SICM and confocal FM}

\begin{figure}
  \includegraphics[width=\textwidth]{gr/correlating/sicm+cfm}
  \caption{%
    \textbf{Combining SICM and confocal fluorescence microscopy.}
    \textbf{a:} Schematic illustration of the correlating data acquired by combining SICM and confocal FM.
    \textbf{b:} 
    	\textit{Left:} Topography of a neuronal network.
    	\textit{Middle:} Fluorescence of the same neuronal network as depicted on the left.
    	\textit{Right:} Overlay of topography and fluorescence of two regions of interest (ROI 1, ROI 2) from
    	the left and middle image.
    Reprinted with permission from ref \cite{Novak2013}.
  }
  \label{fig:sicm+cfm}
\end{figure}

Similar to a SICM and wide-field combination, a combination of SICM and
confocal FM enables imaging of the same region and sample with both techniques
(Fig.~\ref{fig:sicm+cfm}a). The technical implementation involves mounting a
SICM onto an inverted confocal FM. The software used to control the two
microscopes does not have to be the same as both microscopes record data
independently from one another. However, as the recordings are not
simultaneous, pixel-by-pixel correlation is hardly, if at all, possible.

SICM and confocal FM have been combined for the first time by Korchev et al.
\cite{Korchev2000} in 2000. The group used this combination to verify the
volume measurements obtained by SICM. Just a year later, further development
of this combination enabled simultaneous acquisition of Ca\textsuperscript{2+}
imaging and topography data \cite{Shevchuk2001}.
% Afterwards, many different
% studies using a combination of SICM and confocal FM were published:

Seifert et al. \cite{Seifert2017} investigated the morphological activity of
platelets by time-lapse SICM and subsequently obtained confocal recordings of
cytoskeletal proteins of the investigated cells, allowing them to correlate
the activity of the platelets and the structure of the cytoskeleton.
In co-cultures of cardio-myocytes and myofibroblasts, the dynamics of
myocyte-myofibroblast connections was linked to the expression of connexin-43
\cite{Schultz2019}.

SICM has been shown to allow patch-clamp recordings from tiny neuronal
structures like synaptic boutons \cite{Novak2013}. The combination of SICM and
confocal FM allowed to first record an overview image and to identify active
synaptic boutons (Fig.~\ref{fig:sicm+cfm}b), which, in a second step, could be
patch-clamped.

A second study on neurons investigated, besides others, cargo transport in
neuronal dendrites \cite{Takahashi2019}. Here, the combination of
confocal FM and SICM provided a hint that the transport processes observed by
SICM were mediated by mitochondria.

In a study investigating the formation of human immunodeficiency virus-like
particles (VLPs) the correlated SICM-fluorescence confocal microscopy was used
to investigate topography changes together with die distribution of
fluorescently labeled viral proteins like Gag or Viral Protein R
\cite{Bednarska2020}. In contrast to investigating the virus formation with
TIRF microscopy \cite{Jouvenet2008}, the combination of SICM and confocal FM
allowed to investigate the apical side of the cells. It was found that the
release of VLPs is a lot faster than the release at the basal cell side as
known from TIRF investigations and furthermore, a potential impact on the
release time by the application of a fluorescent marker was found
\cite{Bednarska2020}.

% ich würde hier etwas zu potentiellen z-stacks + SICM sagen wollen, das hat
% ja meines Wissens nach noch niemand gemacht...
The applications reviewed above used only a single confocal layer. Yet,
a combination of SICM and $z$-stacking confocal FM could be used to yield more
correlating data. After imaging the entire volume that is scanned by SICM, the
data could be aligned and non-correlating data could be separated from
correlating data. If stage scanning would be omitted, the two measurements
could be performed simultaneously, but not synchronously.

\subsection{Surface scanning}

Scanning surface confocal microscopy (SSCM) \cite{Gorelik2002a} is a
combination of SICM and scanning confocal microscopy (SCM) which enables
simultaneous SICM and FM imaging of the surface of the same sample. In SSCM,
the fluorescence data for each pixel is recorded at the position determined by
SICM for that pixel. Up to now, two slightly different setups have been proposed
(Fig.~\ref{fig:sscm_setup}) \cite{Gorelik2002a,Shevchuk2013}.

\begin{figure}
  \includegraphics[width=\textwidth]{gr/correlating/sscm_setup}

  \caption{%
    \textbf{Two setups for a SSC microscope.}
    \textbf{a:} Diagram of the combined setup with an $x,y,z$ sample scanner.
    \textbf{b:} Diagram of the combined setup with a $x,y$ sample scanner and
    a $z$-probe and objective scanner. 
    Reprinted with permission from ref \cite{Shevchuk2013}.
  }
  \label{fig:sscm_setup}
\end{figure}

In both setups, an SICM is mounted onto an inverted confocal microscope and
before simultaneous recording, the focus of the confocal FM objective is
aligned with the tip of the SICM probe via micrometer screws and the confocal
microscope's focusing mechanics. The
difference between the two setups is the way how the SICM pipette, the
confocal focus and the sample are positioned with respect to each other during
imaging. The
first setup (Fig.~\ref{fig:sscm_setup}a) keeps the 
SICM probe at a fixed position and scanning is performed by the sample
stage which positions the sample in all three dimensions.

In contrast, the second setup (Fig.~\ref{fig:sscm_setup}b) uses the sample
stage to position the sample in $x$- and $y$-direction only while objective
and probe are positioned by additional piezo elements in $z$-direction. Thus,
when the surface is detected by the SICM pipette,  the position of
the confocal spot has to be adjusted by the objective positioning piezo subsequently.

Both setups do not allow to use the controlling software of the confocal
microscope since the fluorescence data has to be recorded after the SICM
pipette has detected the surface. Thus, the SICM controlling soft- and
hardware had to be extended to at least allow reading from the detectors of
the confocal microscope.

Switching from three-axis stage scanning to $z$-movement of the pipette
allowed for faster scan rates as the resonant frequency of the 3D
piezo-actor acted as a limiting factor. Additionally, usage of three
piezo-elements on the same platform led to cross-talk and consequentially
artefacts in the SICM data \cite{Shevchuk2013}.

Both setups keep the probe-objective distance constant and enable the
acquisition of a fluorescence image and a topography of the sample surface in
the same scan. Photobleaching is equalized for every point of the sample
allowing quantitative analysis of the FM data.

\begin{figure}
  \includegraphics[width=\textwidth]{gr/correlating/sicm+scm}
  \caption{%
    \textbf{Combining SICM and SCM.}
    \textbf{a:} Schematic illustration of the correlating data acquired by SSCM.
    \textbf{b:} Simultaneously acquired topographic and fluorescence data of fluorescent Cy3-labeled 
    VLPs adsorbed to the surface of Cos-7 cells. Reprinted with permission from ref \cite{Gorelik2002a}.
    \textbf{c:} SSCM recordings of fixed clathrin-GFP transfected Cos-7 cells. 
    	\textit{Left:} Cell surface topography. 
    	\textit{Center:} Fluorescence recording of the cell shown left. 
    	\textit{Right:} 3D overlay of topographical and fluorescent recordings. 
    Reprinted with permission from ref \cite{Shevchuk2008}.
    \textbf{d:} Live Cos-7 cells sequentially scanned by SSCM. Overlays of simultaneously obtained 
    topographic and fluorescence scans. Fluorescence imaging was started 2\,h 17\,min after the first 
    SICM scan as indicated by the red star. Fluorescent VLPs were added at 2\,h 40\,min. 
    Reprinted with permission from ref \cite{Gorelik2002}.
  }
  \label{fig:sicm+scm}
\end{figure}

As the SICM and SCM data sets are recorded simultaneously, all data points inherently have a
correlating data point in the other set. This immediate correlation, schematically illustrated in
Fig.~\ref{fig:sicm+scm}a, allows not only for correlating studies in fixed
samples (Fig.~\ref{fig:sicm+scm}c), but also studies of dynamics in live
samples (Fig.~\ref{fig:sicm+scm}d).\todo{Improve paragraph}

SSCM has been used for investigating the dynamics of the interaction of
Cy3-labeled polyoma virus-like particles (VLPs) with Cos-7 cells
(Fig.~\ref{fig:sicm+scm}b,d) \cite{Gorelik2002a}, for the investigation of the
uptake of fluorescent nanoparticles in immortalized human alveolar epithelial
type 1 (AT1) cells and AT2 cells \cite{Kemp2008,Novak2014}, for investigating
the differentiation of stem cells \cite{Gorelik2008} and the correlation of
topography and T-tubule distribution in cardiomyocytes \cite{Lyon2012} during
recovery from heart failure. Furthermore, it has been utilized in studies
investigating endocytotic pathways \cite{Shevchuk2008}
(Fig.~\ref{fig:sicm+scm}c) and in a study revealing an alternative closing
mechanism for clathrin-coated pits by recording topography simultaneously with
the fluorescence of tagged actin-binding protein and the tagged clathrin
light-chain \cite{Shevchuk2012}. A further study on clathrin-mediated
endocytosis revealed the effect of mutations in the Dynamin~2 gene on the
maturation and internalisation of clathrin coated pits \cite{Ali2019}.

% Bis hier

% \subsection{Nanoinjection of fluorescently labeled molecules}

% Nevertheless SICM does not necessary being used for investigating the topography of a cell, 
% the nanopipette of a SICMic could be also used for the targeted deposition of molecules, like 
% fluorescently-labeled single stranded-DNA molecules \cite{Ying2002,Hennig2015}. The nanopipette 
% could be also used for nanoinjection of fluorescently labeled molecules into living cells 
% \cite{Hennig2015a}. The SICM was therefore combined with wide-field and direct stochastic 
% optical reconstruction microscopy (dSTORM), and mitochondria, actin and DNA could be successfully 
% stained. One problem with the nanoinjection was that artifacts, due to the holes that were 
% introduced into the membranes, could occur \cite{Hennig2015a}.

% The delivery of fluorescently labeled molecules with the nanopipette could be also used for 
% combining it with single-molecule diffusion investigations \cite{Bruckbauer2007, Bruckbauer2010}. 
% In both studies the scanning ion conductance microscope was combined with TIRF investigations and 
% the diffusion of single molecules over different regions at a spermatozoa were investigated.

% The SICM-nanopipette could be also used for investigating membrane rafts 
% \cite{Simons1997,Pike2006} by coordinated manipulation of focal adhesions. The pipette was 
% functionalized with fibronectin, gangliosides were fluorescently labeled, the pipette deformed the 
% focal adhesion side and the behavior of the fluorescently labeled gangliosides were detected. This 
% was seen as a hint, how membrane rafts reacted \cite{Fuentes2012}.

% Another example for a SICM-nanopipette based manipulation could be found in the study on the 
% generation and distribution of plasma-derived extracellular vesicles (PEVs) on MDA-MB-231 and U87 
% cells. Here the nanopipette were used for both, investigating the topography and for inducing the 
% generation of the PEVs. Additionally the SICM was build up onto a confocal microscope, which could 
% be used for the detection of Calcium \cite{Wang2020}.

% % Piper2006 -> PMID: 17177370
% The SICM-nanopipette could be also used as a nanosensor for pH- or ionic concentration changes 
% \cite{Piper2006}. In this study fluorophores were trapped in the nanopipette and react on changes 
% of the pH value or the concentration of ions with alterations in the detected fluorescence signal.

% Dieser Abschnitt passt nicht so wirklich zu Fluoreszenz-Mikroskopie! Vielleicht noch so was 
% schreiben, wie „SICM kann nicht nur mit Fluoreszenz mikroskopischen Methoden kombiniert werden, 
% sondern kann auch mit … kombiniert werden!? Sonst kann der einfach auch weg. Mal gucken.

% By using a special illumination system consisting of a light emitting diode (LED) light ring, SICM 
% could also being combined with condenser-free phase contrast, darkfield and brightfield images and 
% even Rheinberg illuminations were possible \cite{Webb2014}.

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../manuscript"
%%% End: